File-based localization of numerical perturbations in data analysis pipelines

### 2020-03-24 paper.tex

\toolname makes the following assumptions on the pipeline, meant to ensure that 
a file can be unambiguously associated with the process that created it:
\begin{enumerate}
  \item Files are not written concurrently by different processes, although sequential writes by multiple processes are supported;
  \item In case a process writes multiple versions of the same file, only a single version is made available to other processes.
\end{enumerate} 
In addition, \toolname also requires that processes are associated to a
command line (executable and arguments), to facilitate process instrumentation.

\tristan{Ali: please review these assumptions and check if any other one is required}

@@

\toolname uses the \reprozip tool~\cite{rampin2016reprozip}
to capture: (1) the set of processes created by the
pipeline, and
(2) the set of files read and written by each process, including
temporary files. \reprozip collects this information through the
\texttt{ptrace()} system call, with no required instrumentation of the pipeline.
\toolname then reconstructs a provenance graph by creating process and file
nodes from the \reprozip trace, and by adding directed edges corresponding
to file reads and file writes (Figure~\ref{fig:simple_script}-left). \tristan{the new example is 
confusing: it has too many processes and its purpose is confusing. This should be simplified. 
See suggestion in example.sh: it has only a few processes, and it has a temporary and a multiple-write file. 
If you agree that this example is enough, could you update the figure?}

@@

\begin{figure}
\centering
  \includegraphics[width=1\columnwidth]{images/provenance-graphs}
  \caption{Provenance graphs
  constructed from the example pipeline in
  Listing~\ref{listing:sample-script}.
  Processes are represented with circles and files with rectangles.
  Transient files are shaded in gray. 
  Plain edges 
  represent the process tree (\texttt{fork()} or \texttt{clone()} 
  system calls) \tristan{Does spot make any use of the process tree at any stage? If not, I think we may want to remove these edges.}. Dashed edges represent file dependencies.
  Every node in the graph is labeled using (1) a process id created by our
  reconstruction, (2) the name of the executable run by the process or the file name.
  Process 0 is the initial call to the \texttt{sh} interpreter, and
  the rest of processes are the calls to FSL robustfov, bet, stats and maths made in
  Listing~\ref{listing:sample-script}. Processes 75 and 149 are forked by processes 4 and 78, respectively: 
  they were captured by \reprozip while they did not appear in
  Listing~\ref{listing:sample-script}.  
}
  \label{fig:simple_script}
\end{figure}

@@

As an example, in the sample pipeline in
Listing~\ref{listing:sample-script}, process \texttt{3\#robustfov} writes a
temporary file. Also, file \emph{T1w\_brain.nii.gz} is first written
by \texttt{75\#bet2} and then overwritten by process \texttt{149\#bet2}.

@@

\begin{figure*}
  \centering
  \includegraphics[width=.8\textwidth]{images/labelling-process}
  \caption{Illustration of the labelling approach. Non-reproducible processes are shown in red.}
  \label{fig:iterations}
\end{figure*}

Figure~\ref{fig:iterations} illustrates our incremental labelling 
process for the example in Figure~\ref{fig:simple_script}. 
At step one, process \texttt{3\#robustfov} is executed and then labeled 
as reproducible (green) as it produced files without differences.
At step 2, \texttt{75\#bet2} 
is also labeled as reproducible as it produced files without differences. 
At step 3 \texttt{149\#bet2} is labeled as non-reproducible (red) as it produced files with differences 
from files without differences. Therefore, the files produced by \texttt{149\#bet2} in  
Condition 2 are replaced with the files produced by \texttt{149\#bet2} in 
Condition 1.
At step 4, \texttt{152\#fslmaths} is labeled as non-reproducible as it writes files with differences.
At step 5, \texttt{153\#fslstats} is labeled as 
reproducible as the last process of the pipeline.

>>>>

\toolname makes the following assumptions on the pipeline, meant to ensure that 
a file can be unambiguously associated with the process that created it:
\begin{enumerate}
  \item Files are not written concurrently by different processes, although sequential writes by multiple processes are supported;
  \item In case a process writes multiple versions of the same file, only a single version is made available to other processes.
  \item Temporary files are not removed by the processes that create them.
\end{enumerate} 
In addition, \toolname also requires that processes are associated to a
command line (executable and arguments), to facilitate process instrumentation.

@@

\toolname uses the \reprozip tool~\cite{rampin2016reprozip}
to capture: (1) the set of processes created by the
pipeline, and
(2) the set of files read and written by each process, including
temporary files. \reprozip collects this information through the
\texttt{ptrace()} system call, with no required instrumentation of the pipeline.
\toolname then reconstructs a provenance graph by creating process and file
nodes from the \reprozip trace, and by adding directed edges corresponding
to file reads and file writes (Figure~\ref{fig:simple_script}-left).

@@

\begin{figure}
\centering
  \includegraphics[width=.8\columnwidth]{images/provenance-graphs}
  \caption{Provenance graphs
  constructed from the example pipeline in
  Listing~\ref{listing:sample-script}.
  Processes are represented with circles and files with rectangles.
  Transient files are shaded in gray. 
  Plain edges 
  represent the process tree (\texttt{fork()} or \texttt{clone()} 
  system calls) 
  \tristan{Does spot make any use of the process tree at any stage? If not, 
  I think we may want to remove these edges. 
  Ali: Yes, it uses process tree structure to do some tree operations like search, filter, or remove.
  Also using the process tree we do clustering later}. 
  Dashed edges represent file dependencies.
  Every node in the graph is labeled using (1) a process id created by our
  reconstruction, (2) the name of the executable run by the process or the file name.
  Process 0 is the initial call to the \texttt{sh} interpreter, and
  the rest of processes are the calls to FSL bet, stats and maths made in
  Listing~\ref{listing:sample-script}. Process 73 is forked by process 2: 
  it was captured by \reprozip while it did not appear in
  Listing~\ref{listing:sample-script}.  
}
  \label{fig:simple_script}
\end{figure}

@@

As an example, in the sample pipeline in
Listing~\ref{listing:sample-script}, file \emph{output.nii.gz} is first written
by \texttt{73\#bet2} and then overwritten by process \texttt{75\#fslmaths} as a temporary file.

@@

\begin{figure}
  \centering
  \includegraphics[width=.8\columnwidth]{images/labelling-process}
  \caption{Illustration of the labelling approach. Non-reproducible processes are shown in red.}
  \label{fig:iterations}
\end{figure}

Figure~\ref{fig:iterations} illustrates our incremental labelling 
process for the example in Figure~\ref{fig:simple_script}. 
At step one, process \texttt{73\#bet2} is labeled as non-reproducible (red) 
as it produced files with differences 
from files without differences. Therefore, the files produced by \texttt{73\#bet2} in  
Condition 2 are replaced with the files produced by \texttt{73\#bet2} in 
Condition 1.
At step 2, process \texttt{75\#fslmaths} is executed and then labeled 
as reproducible (green) as it produced files without differences.
At step 3, \texttt{76\#fslstats} is also labeled as 
reproducible as the last process of the pipeline.
